\documentclass{sig-alternate}

\PassOptionsToPackage{hyphens}{url}
\usepackage{hyperref}
\usepackage{color}

\newcommand{\cmw}[1]{{\color{red} CMW: #1}}

\begin{document}

\title{Workshop on Constraints in Software Testing, Verification, and
  Analysis (CSTVA'16)}

\numberofauthors{2}

\author{
  Christoph M. Wintersteiger\\
  \affaddr{Microsoft Research}\\
  \affaddr{21 Station Road}\\
  \affaddr{Cambridge, CB1 2FB, United Kingdom}\\
  \email{cwinter@microsoft.com}\\
  \affaddr{Tel: +44-1223-479-724}\\
  \affaddr{\url{http://research.microsoft.com/people/cwinter}}
  \alignauthor
  Omer Tripp\\
  \affaddr{IBM Research}\\
  \affaddr{1101 Kitchawan Rd}\\
  \affaddr{Yorktown Heights, NY, USA}\\
  \email{otripp@us.ibm.com}\\
  \affaddr{Tel: +1-914-945-1262}\\
  \affaddr{\url{http://researcher.watson.ibm.com/researcher/view.php?person=us-otripp}}
}
\date{October 23, 2013}

\maketitle
\toappear{}

\section{Main contact}

Christoph M. Wintersteiger (See above for contact information.)

\section{Workshop length}

One day

\section{Workshop dates}

Preferably the day before ISSTA, i.e., July 17, 2016. Our second choice is July
21, i.e., the first day after ISSTA ends. 

\section{Workshop Open or Closed}

Open.

\section{Min/Max Number of Participants}

Minimum: 15 and Maximum: 50. We may not organize the workshop if
there are fewer than 15 participants.

\begin{abstract}
During the last decade, there has been significant improvement in the
efficiency and expressive power of Boolean SAT, SMT and Constraint Programming
(CSP/CP) solvers, with a consequent impact on all manner of software
engineering (SE) applications and research programs. A notable example of this
is the rapid development and adoption of solver-based symbolic-execution
techniques for verious purposes, including test generation, security, and
analysis. Despite this increasing use of solvers, there are few venues that are
solely dedicated to bringing together the broader PL/SE and solver communities
in one place. This workshop at a leading PL/SE conference is designed precisely
to fill this gap. The aim of the workshop is to highlight use of solvers in
novel applications, new solver features, and encourage their use in fresh
solutions to long-standing PL/SE problems. Furthermore, as current users become
more sophisticated, they demand richer solver interfaces that are extensible,
tunable and programmable. Designing such interfaces requires deeper interaction 
between solver developers and users. Hence, this workshop will also act as a
venue for feedback from users to developers.

\end{abstract}

\section{Workshop theme and goals}
Recent years have seen increasing interest in the
application of constraint solving techniques to testing, verification
and analysis of software systems. The reason for this interest is the
dramatic improvement in efficiency and expressive power of Boolean
SAT, SMT and CP solvers, thus making it considerably easier to build
and maintain SE applications. A large
number of constraint-based techniques have been proposed and
investigated in model-based testing, code-based testing,
property-oriented testing, statistical testing, equivalence checking,
model checking, fault localization, verification, and program
analysis.  What is even more interesting is that as solvers become
more efficient and expressive, newer applications are being developed
in diverse areas such as security, synthesis, type systems and
software product lines. In the past, solvers were largely used as
a black box. However, more recently a central idea in solver-based SE
applications is deep integration of the solver and the application. The
benefits of deep integration over a black-box approach are many. One such advantage is that the integration can enable
applications to better utilize the adaptive features of modern
solvers. Often the resulting SE techniques are far more powerful and
effective than before.

  It is within this context of increasing use of solvers
in PL/SE, especially through deep integration of extensible solvers with
applications, that we propose to hold the CSTVA workshop at ISSTA, which is
among the leading PL/SE conferences. The goal of the workshop is to bring
together in one place the largely distinct communities of solver and broader
PL/SE researchers. We expect very intense and fruitful cross-community
interaction between the members of the two communities. Such interactions could
lead to completely new approaches to long-standing PL/SE problems, and current
users of solvers can discuss their experience and provide rapid feedback to
solver developers. The workshop could also influence new research directions
into easily extensible, tunable and programmable solvers. Finally, the workshop
can help showcase the breadth and depth of the influence of solvers in SE
research. 

  A prime example of how interaction between solver researchers
and the PL/SE community can lead to impactful research is ``symbolic execution
and its application to testing, verification and analysis''. Symbolic execution
for software testing was proposed by Lori Clarke and J.C. King in the
1970's. However, it didn't really pick up until nearly thirty years later in
the mid-2000's, in large measure due to dramatic improvement in solver
technology. Thanks to SAT solvers like MiniSAT, Lingeling, and ManySAT, and SMT
solvers like Z3, Yices, CVC4, MathSAT, STP, and HAMPI, a variety of
symbolic-execution engines have been developed. Examples of the most successful
such tools include SAGE, PathCrawler, KLEE (and its predecessor EXE), BitBlaze,
WebBlaze, S2E, Java PathFinder and KINT. 

Another example of deep integration of solver and application is
the recently proposed model-checking algorithm IC3. In this approach, not only
is a solver used by the model checker, but also the design of the IC3 algorithm
itself is influenced by ideas from SAT and SMT solving.

A third example comes from the development of specialized CP
techniques for constraints on floating-point numbers, which could not be
accurately resolved using the mathematical tools developed for real
numbers. This is a significant and recent advance in solver technology with
potentially huge impact in automated testing and verification of numerical
software.

We believe that the success we have seen of the combination of
solvers and above-mentioned PL/SE fields can be replicated in other areas of
PL/SE. It is with this goal of increasing the awareness of solvers in the
broader PL/SE community, and increasing interaction between solver and PL/SE
researchers, that the workshop is being proposed.

The workshop will focus on a broad range of topics, where
solvers have already made an impact, e.g. symbolic-execution-based testing, as
well as newer ones where their use is still nascent, e.g. synthesis and
software product lines.

\subsection*{Topics for the Workshop}
\vspace{0.2cm}
\noindent{The topics for the workshop include, but are not limited to,
  the following:}

\begin{itemize}
\item Constraint-based analysis of programs and models
\item Constraint-based test input generation
\item SMT solvers for testing, verification, analysis, and synthesis
\item SMT solvers and applications in computer security
\item Programmable SMT solvers
\item Combinations of constraint solvers
\item Constraint programming and software engineering
\item Solvers and software product lines
\item Solvers and fault localization
\end{itemize}

\section{Relevance to PL/SE research}
The idea of capturing software behavior through logic
constraints is an old but very important idea that has always played a role in
PL/SE.  However, in the past, PL/SE researchers didn't have access to efficient
solvers, and instead had to develop their own solutions to process the
constraints generated by their applications. All that has changed in the last
15 years. Armed with increasingly more powerful SAT/SMT solvers, researchers
now tackle many hard PL/SE problems. As mentioned above, many areas of PL/SE
have been transformed thanks to these methods, and we expect that the broader
PL/SE community will benefit greatly from continued improvement in solver
technology, and deeper integration of solver and application.

\sloppypar{In addition to the above-mentioned applications that
  have been impacted by solvers, new and emerging areas where solvers are used
  heavily include software product lines, constraint-based program analysis and
  fault localization. These application domains are quite distinct from each
  other, and yet they all rely on powerful reasoning engines. For example, in
  software product lines (SPL) research the goal is to show that the
  user-selected features and configuration can lead to a valid product, that
  does not violate important safety and security properties. On the surface,
  the SPL research area seems very different from the problem of fault
  localization, where the problem is to isolate the root cause(s) of an error
  given a program and an error-revealing input. In the constraint-based program
  analysis field, the paradigm is to express the progam analysis problem as
  constraints in declarative languages such as Prolog. These constraints are
  then interpreted by powerful solvers to perform analysis on input
  programs. In the past, researchers in each of these application areas had to
  develop individual specialised solutions to address their problems. As
  already highlighted in the above Section, thanks to the advent of efficient
  and expressive solvers, a new class of systematic methods is being explored
  to solve these knotty problems.} 

\section{Generating discussion}

%% We expect high attendance at the workshop. We plan to organize a
%% panel discussion on the future of solvers in PL/SE that will bring together 2
%% leading solver developers and 2 leading users to present their views on how we
%% can build solvers that better serve PL/SE research needs. 
%% We expect to hold the panel towards the end of the workshop. 
We hope to invite a leading PL/SE researcher to give a keynote address on
his/her experience in using solvers. 

We are also considering the possibility of presentation-only papers (e.g. tool
papers, or papers recently published elsewhere) that can generate considerable
of interest. Such papers can be shorter (say 6 pages), and will not be included
in the proceedings. 

%% \section{Limits on no. of participants}
%% The workshop will not be held if there are fewer than 15 participants
%% and applications to participate will be refused if there are more than
%% 50.

%% \section{Preliminary web site}
%% \url{XXX}

\section{Program Committee}
We are fortunate that many of the leading researchers in the field of SAT/SMT
solvers, CP, PL, and SE applications have served on the program committee (PC)
of CSTVA in the past, and will most likely be happy to serve on the CSTVA PC
also this year. We believe that this reflects the increasing recognition of the
fact that the broader community needs a workshop or conference where developers
of different constraint solving techniques and their users can meet and
exchange ideas. 

The following is a preliminary list of candidates to serve on the program committee: 

\begin{enumerate}
\item Vijay Ganesh, University of Waterloo, Canada
\item Nicky Williams, CEA LIST, France
\item Kapil Vaswani, Microsoft Research, India
\item Aditya Nori, Microsoft Research, India
\item Rupak Majumdar, Max Planck Institute for Software Systems, Germany
\item Koushik Sen, University of California, Berkeley, USA
\item Frank Tip, University of Waterloo, Canada
\item Joxan Jaffar, National University of Singapore, Singapore
\item Nikolaj Bjorner, Microsoft Research, USA
\item Leonardo de Moura, Microsoft Research, USA
\item Cristian Cadar, Imperial College London, UK
\item Arnaud Gotlieb, Simula, Norway and INRIA, France
\item Frederic Dadeau, FEMTO-ST/INRIA, France
\item Krzysztof Czarnecki, University of Waterloo, Canada
\item Julian Dolby, IBM TJ Watson Center, USA
\item Yunhui Zheng, IBM TJ Watson Center, USA
\item Ofer Strichman, Technion, Israel
\item Fran{\c c}ois Bobot, CEA LIST, France
\item S{\'e}bastien Bardin, CEA LIST, France
\item Sylvain Conchon, Universit{\'e} Paris Sud, France
\item Emina Torlak, University of California, Berkeley, USA
\item Daniel Le Berre, Universit{\'e} d'Artois, France
\item Cesare Tinelli, University of Iowa, USA
%% \item Christoph M. Wintersteiger, Microsoft Research, UK
\item Patrick Heymans, University of Namur, Belgium
\item Marsha Chechik, University of Toronto, Canada
\item Xiangyu Zhang, Purdue University, USA
\item Philipp Ruemmer, Uppsala University, Sweden
\item Georg Weissenbacher, Technical University of Vienna, Austria
\item Martin Brain, Oxford University, UK
\item Liana Hadarean, Coverity, USA
\item Ruben Martins, University of Austin, TX, USA
\item Markus N. Rabe, UC Berkeley, CA, USA
\end{enumerate}

\section{Participant solicitation}

Participants will be solicited through a mailshot of the Call For Papers and a
poster. We also intend to use social networks (Facebook, Twitter, etc). This
should reach a wide audience thanks to the diversity of the program committee.
The workshop will be open to all participants until the maximum number of
participants is reached. 

\section{Proceedings}

Accepted papers (except presentation-only papers) will be published in the
workshop proceedings.  The ideal number of papers is around 12, and expected
length is 10 pages with 1-2 additional pages for references in standard ISSTA
format.
\cmw{CEUR? e.g., \url{http://ceur-ws.org/Vol-1163/}, 
Easychair? e.g., \url{http://www.easychair.org/publications/EPiC}}


\section{History of CSTVA}

This proposed workshop will be the 7th of the CSTVA (Constraints in Software
Testing, Verification and Analysis) workshop series. In the recent past, CSTVA
has been held alongside ICST (International Conference on Software Testing) and
ICSE (International Conference on Software Engineering).  

The first CSTVA meeting was held alongside the CP conference (Principles and
Practice of Constraint Programming) in Nantes, France, in 2006 and attracted
more than 25 participants. It also featured an invited presentation by Andy
King (University of Kent).
See \url{http://www.irisa.fr/manifestations/2006/CSTVA06/}.

The second edition took place in conjunction with the ICST 2010 in Paris,
France, attracting more than 30 participants. It also featured an invited
presentation by Peli de Halleux (senior researcher at Microsoft Research).  
See \url{http://www.st.cs.uni-saarland.de/cstva10}. 

A third meeting at ICST 2011 in Berlin, Germany, also attracted a similar
number of attendees, and featured an invited presentation of Patrice Godefroid
(principal researcher at Microsoft Research).  
See \url{http://www.st.cs.uni-saarland.de/cstva11}.

A fourth edition was held at ICST 2012 in Montreal, Canada which attracted
close to 40 participants. It also featured a keynote presentation by Vijay
Ganesh (assistant professor at the University of Waterloo, and a co-organizer
of the proposed workshop.) 
See \url{http://srg.doc.ic.ac.uk/cstva12/}.

A fifth edition was held at ICST 2013 in Luxembourg which attracted more than
40 participants. It featured a keynote address by Sarfraz Kurshid (University
of Texas, Austin). 
See \url{http://cstva2013.univ-fcomte.fr/index.php?aim}.

A sixth edition was held at ICSE 2014 in Hyderabad, India, which attracted a
similar number of participants participants. It featured a keynote address by
Nikolaj Bjorner (Microsoft Research). 
See \url{http://srg.doc.ic.ac.uk/cstva14/}.

\cmw{Necessary?}
A seventh edition was planned to be held at ISSTA 2015, but it did not attract
enough participants and had to be cancelled. We plan to take precautions
against a recurrence of this by advertising the workshop more widely and by
soliciting submissions from leading figures in the relevant areas directly.
See \url{http://www.cs.utah.edu/~rsas/cstva15/}.

Given the success of previous editions of this workshop, we believe that the
current version that we are proposing will also be successful. We believe that
by co-locating with ISSTA and broadening the scope of the workshop to go beyond
testing, verification and analysis to also include synthesis, security,
software product lines, etc, we will increase the chances of success of the
workshop dramatically. We are hopeful that we will attract more than 40
participants. 

\section{Logistic constraints}

Wifi connectivity for the participants and on-line demonstrations.

\section{Organizer background}

\begin{enumerate}
\item {\bf Christoph M. Wintersteiger} is a researcher at Microsoft Research and
  one of the authors of the Z3 SMT solver and theorem prover. He works on
  decision procedures for finite domains, e.g., for SAT, QBF, or bit-vectors
  and floating-point numbers as in SMT (QF\_)BV and (QF\_)FP as well as
  applications thereof in software/hardware verification and in newly emerging
  application areas like computational biology.  

  Christoph has (co-)organized workshops (e.g., SMT) before and he frequently
  serves on the PC of closely related conferences and workshops, as well as the
  SMT steering Committee. 

\item {\bf Omer Tripp} is a leading researcher in the broad area of programming
  languages at the IBM TJ Watson Research Center. Omer has worked both on SMT
  solvers, and in particular string solvers, and on applications of SMT
  solvers, including concurrency analysis, security analysis and runtime
  enforcement of security and privacy features. 
  
  Omer has organized, or served as a member of the program committee of,
  several relevant workshops and conferences. These include PLAS 2014 and ACSAC
  2016. Given his intensive work recently on solvers and applications thereof,
  this year Omer has decided to participate for the first time in the
  organisation of CSTVA. 
\end{enumerate}

\end{document}
