%% \chapter*{Preface}
\section*{Preface}

This Conference Proceedings volume contains the written versions of the
contributions presented during the Seventh International Workshop on Constraint
Solvers in Testing, Verification and Analysis (CSTVA'16). This year the
workshop was co-located with the International Symposium on Software Testing
and Analysis (ISSTA). The venue was Saarbr{\"u}cken, the capital of the
Saarland, which is the smallest German federal state. It is located near the
French border, halfway between Paris and Frankfurt on the high-speed railway
that connects these two cities in less than four hours. Saarbr{\"u}cken offers
picturesque attractions and places of historic interest, making it a lovely
destination for a hike or a day trip.  

As in previous years, this year too CSTVA provided a platform for researchers
and practitioners from the solver community on the one hand, and the
programming-languages (PL) and software-engineering (SE) communities on the
other hand, to interact, share feedback, exchange views, and form an ambitious
and unified vision as to the role of solvers in the PL and SE research
domains. The importance of this discussion stems from the rapid improvement in
expressive power of Boolean satisfiability (SAT),
satisfiability-modulo-theories (SMT) and constraint-programming (CP) solvers
throughout the last decade. These impressive advances enable the integration of
such solvers into PL/SE techniques and algorithms in new and powerful ways.
This year, the workshop covered a wide variety of topics, including use of
solvers to reason about weak memory models, multi-threaded software, to
implement concolic as well as combinatorial testing efficiently, perform
bounded model checking, and the properties of industrial SAT solving problems.
%% and *invited talks*. 
%% It attracted participants from the United States, Germany, *more places*. 
The program consisted of three types of contributions: short papers, full
papers and invited talks. There were two full-paper presentations, two
short-paper presentations and two keynote talks. 

We would like to thank all the participants for their contributions to the
workshop program as well as for their contributions to these Proceedings. We
extend special thanks to the German participants for their support and
hospitality, which allowed all foreign participants to feel more at home. We
are also grateful to the invited speakers --- Vijay Ganesh from the University
of and Murali Krishna Ramanathan from the Indian Institute of Science (IIS),
Bangalore --- for their inspiring talks and unwavering acceptance of our
invitation.

We look forward to the Eighth International Workshop on Constraint Solvers in
Testing, Verification and Analysis. We hope that it will be as interesting and
enjoyable as previous CSTVA meetings. 

~\bigskip

\noindent
\begin{minipage}[t]{.4\textwidth}
July, 2016
\end{minipage}%
\hfill
\begin{minipage}[t]{.4\textwidth}\flushright
Omer Tripp\\
Christoph M. Wintersteiger
\end{minipage}
